Title Goes Here

### 2020-02-25 paper/paper.tex
Improved wording in description of marking
Slight change to where marks are removed and add corresponding transformation names to flowchart

If an expression already has an EDSL type (a type of the form \ttt{E a} for some
\ttt{a}), then it does the marking procedure ignores it and does not wrap it with
\ttt{externalize}.

@@

\begin{tikzpicture}[node distance=2cm]
\node (start) [startstop] {Core from GHC};
  \node (tailrec) [process, below of=start]{Transform tail recursion of definitions that have a mark (when applicable)};
  \node (findmark) [process, below of=tailrec]{Find marked expression and erase mark};
    \node (handlecase) [process, below of=findmark]{Mark scrutinee};
    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{Encode as EDSL primitive};
    \node (handleapp) [process, right of=handleprim, xshift=1.6cm]{Wrap with \ttt{Construct} and \ttt{ConstructAp}};

    \node (marksubs1) [process, below of=handleprim]{Mark sub-expressions};
    \node (marksubs2) [process, below of=handleapp]{Mark sub-expressions};

    \node (onebranch) [process, below of=handlecase]{Transform one branch};
    \node (prodmatch) [process, below of=onebranch]{Transform one field in product match of current branch};
    \node (abstractvar) [process, below of=prodmatch]{Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
    \node (markbody) [process, below of=abstractvar]{Mark branch body};
    \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};

@@

\begin{align*}
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
    &:=
        \ttt{\{ K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
    &:=
      \begin{cases}
        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body}}\\
        \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}
      \end{cases}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
  T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
    &:=
      \begin{cases}
        \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
        \ttt{Done (case s of \{ ... \})} & \text{otherwise}
      \end{cases}
\end{align*}
\begin{align*}
  &C^{*}\expr{\ttt{x}} :=
    \begin{cases}
      \ttt{x} & \text{if \ttt{x} has no subexpressions marked with \ttt{externalize}}\\
      C^{*}\expr{C\expr{\ttt{externalize y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\
    \end{cases}\\
\end{align*}
\begin{align*}
  &C\expr{\ttt{externalize (runIter x)}} := \ttt{TailRec $M\expr{\ttt{x}}$}\\
  &C\expr{\ttt{externalize (x + y)}} := \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
  &C\expr{\ttt{externalize (case scrutinee of \{ ... \})}} := \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
  &C\expr{\ttt{externalize ($\lambda$x $\rarr$ body)}} := A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
  &C\expr{\ttt{externalize (f x)}} := \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
  &C\expr{\ttt{externalize x}} := \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
  &\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} := \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} := \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
  &B\expr{\ttt{\{ \}}} := \ttt{EmptyMatch}\\
  &\\
  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} := \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
  &P\expr{\ttt{x $\rarr$ body}} := \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
  &P\expr{\ttt{ $\rarr$ body}} := \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
  &\\
  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} := \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
  &\\
  &M\expr{\ttt{(x :: a)}} :=
    \begin{cases}
      \ttt{externalize x} & \text{if}\; \nexists t, a\; \typeeq E\; t\\
      \ttt{x} & \text{if}\; \exists t, a\; \typeeq E\; t
    \end{cases}
\end{align*}

>>>>

If an expression already has an EDSL type (a type of the form \ttt{E a} for some
\ttt{a}), then the marking procedure ignores it and does not wrap it with
\ttt{externalize} (see $M$ in Section 7.2).

@@

In the following flowchart, the names in parentheses in a node refers to the names of transformations from the following section (Section 7.2).

\begin{tikzpicture}[node distance=2cm]
\node (start) [startstop] {Core from GHC};
  \node (tailrec) [process, below of=start, yshift=-0.5cm]{($D, T, T'$) Transform tail recursion of definitions that have a mark (when applicable)};
  \node (findmark) [process, below of=tailrec, yshift=-0.5cm]{($C^*$) Find marked expression and erase mark};
    \node (handlecase) [process, below of=findmark]{($C$) Mark scrutinee};
    \node (handleprim) [process, right of=handlecase, xshift=1.6cm]{($C$) Encode as EDSL primitive};
    \node (handleapp) [process, right of=handleprim, xshift=1.6cm]{($C$) Wrap with \ttt{Construct} and \ttt{ConstructAp}};

    \node (marksubs1) [process, below of=handleprim]{($C$) Mark sub-expressions};
    \node (marksubs2) [process, below of=handleapp]{($C$) Mark sub-expressions};

    \node (onebranch) [process, below of=handlecase]{($B$) Transform one branch};
    \node (prodmatch) [process, below of=onebranch]{($P$) Transform one field in product match of current branch};
    \node (abstractvar) [process, below of=prodmatch]{($P$) Abstract the pattern variable of the field with \ttt{Lam} and \ttt{Var}};
    \node (markbody) [process, below of=abstractvar]{($P$) Mark branch body};
    \node (endstep) [process, below of=markbody, right of=markbody, xshift=1.6cm]{End of one transformation step};

@@

\begin{align*}
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ case s of \{ ...$_A$ \}; ...$_B$ \}}}\\
    &:=
        \ttt{\{ K x0 ... xN $\rarr$ $T_{\ttt{f}}\expr{\ttt{case s of \{ ...$_A$ \}}}$; $T'_{\ttt{f}}\expr{\ttt{\{ ...$_B$ \}}}$\}}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}}\\
    &:=
      \begin{cases}
        \ttt{\{ K x0 ... xN $\rarr$ body}[\ttt{f} \mapsto \ttt{Step}]\ttt{; }T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}\ttt{ \}} & \text{if \ttt{f} occurs free in \ttt{body}}\\
        \ttt{\{ K x0 ... xN $\rarr$ Done body; ... \}} & \text{otherwise}
      \end{cases}\\
  T'_{\ttt{f}}&\expr{\ttt{\{ \}}} := \ttt{\{ \}}\\
  T_{\ttt{f}}&\expr{\ttt{case s of \{ ... \}}}\\
    &:=
      \begin{cases}
        \ttt{case s of $T'_{\ttt{f}}\expr{\ttt{\{ ... \}}}$} & \text{if \ttt{f} occurs free in \ttt{\{ ... \}}}\\
        \ttt{Done (case s of \{ ... \})} & \text{otherwise}
      \end{cases}
\end{align*}
\begin{align*}
  &C^{*}\expr{\ttt{x}} :=
    \begin{cases}
      \ttt{x} & \text{if \ttt{x} has no subexpressions marked with \ttt{externalize}}\\
      C^{*}\expr{C\expr{\ttt{y}}} & \text{if \ttt{externalize y} is the first marked subexpression of \ttt{x}}\\
    \end{cases}\\
\end{align*}
\begin{align*}
  &C\expr{\ttt{runIter x}} := \ttt{TailRec $M\expr{\ttt{x}}$}\\
  &C\expr{\ttt{x + y}} := \ttt{Add $M\expr{\ttt{x}}$ $M\expr{\ttt{y}}$}\\
  &C\expr{\ttt{case scrutinee of \{ ... \}}} := \ttt{CaseExp $M\expr{\ttt{scrutinee}}$ } B\expr{\ttt{\{ ... \}}}\\
  &C\expr{\ttt{($\lambda$x $\rarr$ body)}} := A\expr{\lambda \ttt{x} \rarr \ttt{body}}\\
  &C\expr{\ttt{(f x)}} := \ttt{ConstructAp $M\expr{\ttt{f}}$ $M\expr{\ttt{x}}$}\;\;\;\;\text{(Where \ttt{f} is not a primitive or an application of a primitive)}\\
  &C\expr{\ttt{x}} := \ttt{Construct x}\;\;\;\;\text{(Where \ttt{x} is a variable which is not a primitive)}\\
  &\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body0; ... \}}} := \ttt{SumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body0}}\; B\expr{\ttt{\{ ... \}}}\\
  &B\expr{\ttt{\{ K x0 ... xN $\rarr$ body \}}} := \ttt{OneSumMatchExp }P\expr{\ttt{x0 ... xN $\rarr$ body}}\\
  &B\expr{\ttt{\{ \}}} := \ttt{EmptyMatch}\\
  &\\
  &P\expr{\ttt{x0 x1 ... $\rarr$ body}} := \ttt{ProdMatchExp }A\expr{\ttt{$\lambda$ x0 $\rarr$ $P\expr{\ttt{x1 ... $\rarr$ body}}$}}\\
  &P\expr{\ttt{x $\rarr$ body}} := \ttt{OneProdMatchExp }A\expr{\ttt{$\lambda$ x $\rarr$ body}}\\
  &P\expr{\ttt{ $\rarr$ body}} := \ttt{NullaryMatch $M\expr{\ttt{body}}$}\\
  &\\
  &A\expr{\ttt{$\lambda$(x :: a) $\rarr$ body}} := \ttt{Lam (Name @a uniq) $M\expr{\ttt{body$[\ttt{x} \mapsto \ttt{Var uniq}]$}}$}\text{\;\;\;\;\;(where \ttt{uniq} is a globally unique identifier)}\\
  &\\
  &M\expr{\ttt{(x :: a)}} :=
    \begin{cases}
      \ttt{externalize x} & \text{if}\; \nexists t, a\; \typeeq E\; t\\
      \ttt{x} & \text{if}\; \exists t, a\; \typeeq E\; t
    \end{cases}
\end{align*}
